---
title: "Beamer Presentation - HW3 BIO 381"
author: "Adrian Wiegman"
date: "February 3, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Primary objective of the Presentation

- demonstrate use of sequential and simultaneous bullet points
- use colored text
- use an image 
- use an image with a caption
- use an image with an overlayed caption

## Secondary objective of the presentation

- talk about my what I had for breakfast!

## Ingredients of my breakfast 

>- \color{yellow} eggs
>- \color{magenta} ham
>- \color{green} green dye
>- \color{black} butter

## An equation for the weight I gained after eating my breakfast

$Q = k(E*n_E + GD + H*n_H + B)$


- *Q* is the mass of my breakfast (g)
- *E* the mass of on large chicken egg (g)
- *n~i~* is the number of i^th^ ingredient  
- *GD* is the mass of green dye added to the egg yolk prior to frying
- *H* the mass of a standard 1 cm thick slice of ham shank with a diameter of 15cm (g)
- *B* the mass of butter added, to greese the pan 
- *k* k~ce~ is the moisture loss coefficient (g g^-1^) to heat all ingredients in a frying pan from 5^o^C to a mean tempurature of 70^o^C

## An R vector of what I had for breakfast

```{r cars, echo = TRUE}
library (ggplot2)
ingredients<- c('greenDye','eggs','ham','butter')
mass <- c(0.05,70,50,5)
names(mass) <- ingredients
k <- 0.8
n <- c(1,1,2,1)
names(n) <- ingredients
Q <-sum(n*k*(mass))
print(Q)
```

## A picture of what I had for breakfast
 \begin{picture}(320,250)
 \put(85,75){\includegraphics[height=6cm]
 {samiam.jpg}}
 \put(00,60){\begin{minipage}[t]{1\linewidth}
 \end{minipage}}
 \end{picture}

## A properly annotated picture of what I had for breakfast
\begin{picture}(320,250)
 \put(85,75){\includegraphics[height=6cm]
 {samiam.jpg}}
 \put(00,60){\begin{minipage}[t]{1\linewidth}
 green eggs and ham
 
{\color{blue}(source: http://www.drseussart.com/illustration-art/green-eggs-and-ham-50th-anniversary-print)}%% caption
 \end{minipage}}
 \end{picture}





